---
title: "Data607 - Project 1: Chess Tournament Results"
author: "Anthony Josue Roman"
date: "2024-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(tidyverse)
```

This project will use the following packages:

```{r eval=FALSE}
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(tidyverse)
```

## General Overview

This file will be available on my [GitHub Page](https://github.com/spacerome/Data607_Project1).

This project will primarily focus on manipulating a dataset from a `.txt` file named `tournamentinfo.txt`. This `rmd` file will create a `.csv` file after analyzing the data within the file.

## Chess Tournament Results Extraction

Initially we will extract the following data from the `.txt` file to prepare to analyze the data before exporting it into a `.csv` file.

```{r echo=TRUE}

a <- getwd() # Just to set directory
setwd(a)

# Now use this function to load the data
tournamentData <- suppressWarnings(readLines("tournamentinfo.txt"))

# Preview the data after removing the first four lines
head(tournamentData)

```
## Initializations

The following code blocks will be using the following data from  `tournamentData` and converting it into two matrices for manipulation, `d1` will consist of the names and `d2` will consist of the other values which we will use later on.

```{r echo = TRUE}
tdm <- matrix(unlist(tournamentData), byrow = TRUE)

d1 <- tdm[seq(5, length(tdm),3)]

head(d1)
```

```{r echo = TRUE}
d2 <- tdm[seq(6, length(tdm),3)]

head(d2)
```
## String Manipulation

The following code block will modify the following strings for us to utilize the following variables `playerName` for player name, `playerState` for the state of the player, `totalPoints` for the total points of the player, `preRating` for the pre-rating, and `numRounds` for the number of rounds. This will be utilized to help us initialize the data to calculate the Average rating of the player's opponents. The following matrices `d1` and `d2` will be utilized here for the values for each variable such as `playerName`, `playerState`, `totalPoints`, `preRating`, and `numRounds`.

```{r echo = TRUE}
playerNum <- as.numeric(str_extract(d1, '\\d+'))
pName <- str_extract(d1, '[A-z].{1,32}')
playerName <- str_trim(str_extract(pName, '.+\\s{2,}'))
playerState <- str_extract(d2, '[A-Z]{2}')
totalPoints <- as.numeric(str_extract(d1, '\\d+\\.\\d'))
preRatingraw <- str_extract(d2,'R:.{8,}-')
preRating <- as.numeric(str_extract(preRatingraw, '\\d{1,4}'))
numRoundsraw <- suppressWarnings(str_extract_all(d1,'[A-Z]\\s{2,}\\d+'))
numRounds <- suppressWarnings(str_extract_all(numRoundsraw, '\\d+'))
```

## Average Chess Opponent Ratings

The following code block here will calculate the average chess opponent ratings for each player by utilizing a for loop where it calculates the mean value for each player. The value will be stored in `avg_chess_opp_rating` where `i` is the location for each player. 

```{r echo = TRUE}
avg_chess_opp_rating <- c()

for(i in c(1:length(numRounds))){
  avg_chess_opp_rating[i] <- round(mean(preRating[as.numeric(numRounds[[i]])]),0)
}

head(avg_chess_opp_rating)
```

## Modified Tournament Data

After getting the ratings, the following values are then exported into `modifiedtournamentData` to be prepared to be exported into a csv file. The following values will be stored into a data frame: `playerName`, `playerState`, `totalPoints`, `preRating`, and `avg_chess_opp_rating`. 

```{r echo = TRUE}
modifiedtournamentData <- data.frame(playerName, playerState, totalPoints, preRating, avg_chess_opp_rating)

head(modifiedtournamentData)
```

## Export Data to CSV

Lastly, the code block will export the values into a `.csv` file named `new_tournament_info.csv`. 

```{r echo = TRUE}
write.csv(modifiedtournamentData, file = "new_tournament_info.csv")
```

## Conclusion

Overall, this project was not that bad, I did attempt utilizing a function, which kept failing and kept outputing its values incorrectly after the first part, so I made it simpler. To make the function work, I will continue to troubleshoot it to make it perform better, and it will be included in my GitHub Page.